Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-18 thesis/chapters/ch_introduction.tex
Thesis intro updates

Dwarf galaxies have been called the building blocks of the Universe, and represent some of the best laboratories to study the fundamental physics of galactic evolution. From a theoretical perspective, their small physical size, relatively quiet accretion history, and low star formation rates make simulating their evolution in detail at high resolution substantially less computationally intensive than more massive galaxies like the Milky Way.

\textit{Mention study of dwarf galaxies instead of MW first as way to simulate physics in detail and build up undserstanding to more massive regimes. Lead into physics that is important below.}

@@

\subsection{Nucleosynthesis and Stellar Evolution} \label{intro:sec:nucleosynthesis}

@@

While there is no real sharp transition where a galaxy is considered a ``dwarf" galaxy, historical definitions are based on luminosity, using the SMC and LMC as rough anchor points for the upper-end of the dwarf galaxy scale. \cite{Bullock2017} suggests a deliniation based on stellar mass. For this dissertation, we follow this definition and consider galaxies with $M_* < 10^9$~M$_{\odot}$ as dwarf galaxies, and galaxies with $M_* < 10^5$~M$_{\odot}$ as UFDs.

@@

Until recently, the study of chemical evolution in Local Group dwarf galaxies has focused predominantly on the stellar component. This is due in large part because, though dwarf galaxies have higher gas fractions than more massive galaxies \citep{Geha2006}, many of these galaxies have undergone environmental stripping processes, rendering them devoid of a measurable gas component \citep{GrcevichPutman2009}.

% mention difficulties of measuring gas phase abundances in dwarf galaxies?

\subsection{Metals in the ISM and CGM of Dwarf Galaxies}

The proximity of Milky Way dwarf galaxy satellites allows us to make some of the most detailed extragalactic stellar abundances measurements.

Stellar feedback drives outflows, galactic winds, from galaxies across all scales (references). Dwarf galaxies in particular, with their relatively shallow potential wells, drive effective galactic winds, with mass loading factors ($\eta$ = $\dot{\rm{M}}_{\rm out} / {\rm SFR}$) ranging from XXX to XXX (cite, muratov etc.). How exactly stellar feedback couples to the ISM to drive these stellar winds is still uncertain, and is the topic of substantial ongoing research, but effective winds

@@

>>>>

Dwarf galaxies have been called the building blocks of the Universe, and represent some of the best laboratories to study the fundamental physics of galactic evolution. From a theoretical perspective, their small physical size, relatively quiet accretion history, and low star formation rates make simulating their evolution in detail at high resolution substantially less computationally intensive than more massive galaxies like the Milky Way. Although their lower brightness limits the number of dwarf galaxies that we can observe, the abundance and relative proximity of dwarf galaxies in the Local Group allows for a significant sample of galaxies with resolved stellar populations. Observations of these resolved stellar populations can be used to derive detailed star formation histories and chemical abundances that can be used together to test our theoretical understanding of galactic chemical evolution in a variety of contexts.

\textbf{need final par}

@@

\subsection{Nucleosynthesis} \label{intro:sec:nucleosynthesis}

@@

While there is no real sharp transition where a galaxy is considered a ``dwarf" galaxy, historical definitions are based on luminosity, using the SMC and LMC as rough anchor points for the upper-end of the dwarf galaxy scale. \cite{Bullock2017} suggests a deliniation based on stellar mass. For this dissertation, we follow this definition and consider galaxies with $M_* < 10^9$~M$_{\odot}$ as dwarf galaxies, and galaxies with $M_* < 10^5$~M$_{\odot}$ as UFDs. In general, dwarf galaxies are extremely dark matter dominated, with mass to light ratios over 10$^2$, \citep{SimonGeha2007,Strigari2009,Wolf2010}, convert gas into stars at a lower efficiency than more massive galaxies (cite), and, for the lowest mass dwarf galaxies, exhibit bursty star formation histories driven by stochasticity in the process of both star formation and the effects of stellar feedback (cite).

@@

The first studies of stellar abundances in the Local Group focused on stars in nearby dSph's, and massive stars in the LMC, SMC, and similar, more massive dwarf galaxies \citep[e.g.][]{}. These studies typically obtained abundances for only a small sample ($\sim$ 10) stars, limiting the extent to which they can be used to test models of galacitc chemical evolution.  However, the advent of high-resolution spectrographs at both Keck and the Very Large Telescope -- capable of obtaining the spectra of many stars with a single pointing -- have dramatically increased the sample size of stellar abundances in nearby dwarf galaxies \citep[e.g.][]{}.

Stellar feedback drives outflows, galactic winds, from galaxies across all scales (references). Dwarf galaxies in particular, with their relatively shallow potential wells, are expected to drive effective galactic winds \citep{MacLowFerrara1999} with mass loading factors ($\eta$ = $\dot{\rm{M}}_{\rm out} / {\rm SFR}$) above XXX \citep{Muratov2015,Christensen2018}. Not only does this correspond to a significant amount of mass outflow from these galaxies, but these winds should also carry out a substantial amount of metals from these galaxies. \cite{Kirby2011-metals} used measurements of the stellar abundances in Local Group dSph's and their resolved star formation histories to determine that their stars contain less than 5\% of the metals synthesized in each galaxy. However, since these galaxies have lost their gas through environmental processes as satellites of the Milky Way (ram pressure and tidal stripping), it was unclear if this loss could be attributed entirely to feedback. However, using gas and stellar abundances \cite{McQuinn2015} showed that Leo P -- a gas-rich Local Group dwarf galaxy -- has also ejected $\sim$95\% of its metals. This is significant evidence for metal-enriched, feedback-driven galactic winds operating in these low mass dwarf galaxies.

Finally, the low metallicities of the lowest mass dwarf galaxies make them ideal places to better constrain large uncertainties in stellar nucleosynthesis \citep{FrebelNorris2015,Frebel2018}. At these low metallicities, the number of possible enrichment events that could have possibly contributed to the observed stellar abundances decreases significantly, increasing the possibility of using these abundance patterns to constrain the yields of individual sources. This is particularly true for exotic enrichment events, like those expected to be significant sources of r-process elements, which are potentially so rare that only one or fewer events would be expected in an individual low-mass galaxy.

The most prominant example of these types of studies has been of the UFD Reticulum II, where \cite{Ji2016} used the extreme r-process enhancement in its stars to argue in favor of a neutron star -  neutron star origin for the enrichment. This conclusion has been supported by observations in additional Local Group dwarf galaxies \citep{Duggan2018}, but there is still some disagreement as to this conclusion \citep[e.g.][]{Siegel2018}. Interestingly, the lack of these extreme r-process abundances in other UFDs suggests that another source of r-process must exist \citep{Ji2016b}. Some of the significant uncertainties in interpreting these observations lie with understanding how r-process elements mix within the ISM and are ejected from low mass dwarf galaxies. Recent research has examined how metals from individual enrichement events may mix and evolve in a variety of contexts \citep[e.g.][]{Bland-Hawthorn2015,Ritter2015,Montes2016,Safarzadeh2017}, but there is still much to be understood about this process. In particular, there is still much to be gained from adapting insights from these types of simulations to the modelling used to interpret observations of stellar abundances (as discussed in more detail in Section~\ref{intro:sec:onezone}).

@@

\textbf{Include a figure? Maybe from recent spectra paper? Or HI from ALFALFA? or both?}
